IT Spillovers in Long-Run TFP \\ A SVAR Approach

### 2018-04-26 Latex/Slides_Presenter_April2018/dw23jan2018.tex
Preparing slides for the macro reading presenter

\date{January 23, 2018}

@@

	\begin{itemize}
	\item The neoclassical tradition (RBC literature) in macro has treated total factor productivity (TFP) as exogenous, working against endogenous growth models

	(King \& Rebelo 1999, Christiano, Eichenbaum \& Trabandt 2014, 2018, vs. e.g. Romer 1990)

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation II}

	\begin{itemize}
	\item Our main analysis is estimating a structural VAR in which we identify a shock to the exogenous component of future TFP (a news shock) and a shock to the endogenous component of TFP (a shock to IT productivity, hereafter ``IT shock").

@@

	\item Along the way, we provide a solution to an econometric challenge in the literature on long-run productivity.

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Our starting point: Barsky \& Sims (2011)}
	\label{BS_quote}

BS identify news shocks as shocks that maximize future fluctuations in TFP. But they warn:

\blockquote{\emph{A more general objection to our empirical approach would be that a number of structural shocks, which are not really ``news'' in the sense defined by the literature, might affect a measure of TFP in the future without impacting it immediately. Among these shocks might be research and development shocks, investment specific shocks, and reallocative shocks. Our identification (and any other existing VAR identifications) would obviously confound any true news shock with these shocks.}}

\hspace{5cm} Barsky \& Sims (2011), p. 278.

$\rightarrow$ We focus on a specific subset of shocks which are ``not really `news''': shocks to the productivity of the sector that produces IT goods.

%\hyperlink{BS_FEV}{\beamerbutton{Barsky \& Sims' FEV-table}}	
\end{frame}
%%%%%%%%%%%%%%%%%

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A toy model}

	How should we think of this IT productivity shock?

	Contextualize using a simple expanding variety model (Romer 1990)

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Expanding variety model}
	\label{expanding_variety}

Final good $Y_t$ is the CES aggregate of a number $A_t$ intermediate goods $Y_t^M(s)$:	% theta is between 0 and infinity
	\begin{equation}
	Y_t = \Big[ \int_{0}^{A_t} Y_t^M(s)^{\frac{\theta-1}{\theta} } ds \Big]^{\frac{\theta}{\theta-1}}
	\end{equation}

\pause

Intermediate good $s$ is Cobb-Douglas with exogenous TFP $\Psi_t$:
\begin{equation}
Y^M_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
\end{equation}

\pause

$\rightarrow$ letting $K_t$ and $L_t$ be aggregate capital and labor, final output is:
\begin{equation}
Y_t = A_t^{\frac{1}{\theta-1}} \Psi_t K_t^{\alpha}L_t^{1-\alpha}
\end{equation}

\pause

IT sector is a separate sector that uses IT investment and its own sectoral productivity $\lambda_t$ to expand $A_t$.  \hyperlink{it_sector}{\beamergotobutton{IT sector in detail}}

$\rightarrow$ IT sector is driven by profits from selling the IT good, but its advances spill over to aggregate TFP.

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{The shocks}

\begin{equation*}
Y_t \; = \underbrace{A_t^{\frac{1}{\theta-1}}}_\text{endogenous TFP} \; \; \underbrace{\Psi_t}_\text{exogenous TFP} \; K_t^{\alpha}L_t^{1-\alpha}
\end{equation*}

$\hookrightarrow$ A news shock is $\Psi_{t+k} \uparrow$ for some $k>0$

$\hookrightarrow$ An IT productivity shock is $\lambda_t \uparrow$, which shows up in $A_{t+j} \uparrow$ with a lag $j>0$

So in this model, the innovation to TFP at period $t$ is:
\begin{equation}
\epsilon^{TFP}_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{f(IT_{t-j})}_\text{IT productivity shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation}

Crucially, both lead to an increase in future TFP, with no effect on impact.

@@

Intuition:
\begin{itemize}
\item News shock makes all sectors of the economy more productive $\rightarrow$ impacts the sectoral price $P^{IT}$ as well as P $\Rightarrow$ relative prices do not move.
\item IT productivity shock makes only the IT sector more productive $\rightarrow$ impacts the sectoral price $P^{IT}$ \textbf{only} $\Rightarrow$ relative prices move! 
\end{itemize}

>>>>

\date{April 26, 2018}

@@

@@

	\item However, an increase in IT investment affects TFP with some delays. Therefore this shock is likely to look very similar to a standard news shock.

@@

	\begin{itemize}
	\item Our main analysis is estimating a structural VAR in which we identify a shock to the exogenous component of future TFP (hereafter \textit{news shock}) and a contemporaneous shock which incentivizes investment in information and technology stock (hereafter \textit{IT shock}).

@@

    \item As a support for our identification we design a 2-sector general equilibrium model which features both investment in IT and news shocks.

%%%%%%% Slide %%%%%%
\begin{frame}
\frametitle{Preview of the Results}

\begin{itemize}
	\item IT shocks drive about half of long-run TFP fluctuations in our sample.

	\item News shocks drive approximately one fifth $\Rightarrow$ Still relevant but - once appropriately cleaned from IT shocks - less important then found previously in the literature.

	\item Our structural model points out that relative prices can be used to disentangle the two.

@@

@@

@@

Intuition:
\begin{itemize}
\item As long as the technology improvement heralded by the news shock does not realize, relative prices should not move.
\item An increase in IT productivity shock makes only the IT sector more productive $\rightarrow$ impacts the sectoral price $P^{IT}$ \textbf{only} $\Rightarrow$ relative prices move! 
\end{itemize}
